### 2016-04-11 ms/ms_probaweb.tex

10: University of Canterbury at Christchurch, School of Biological Sciences\\

@@

Biogeography has historically focused on the spatial distribution and
abundance of species, neglecting variation in the way species interact with
each other. Models of species distribution and of interactions have
historically taken different paths. Here, we plea for an integrated approach,
adopting the view that community structure is best represented as an
ecological network of interactions. We outline a conceptual approach,
suggesting that the ecological niche may be redefined to encompass the effect
of the environment on species distribution (the Grinnellian dimension of the
niche) and on the ecological interactions among them (the Eltonian dimension).
Starting from this novel concept, we develop a quantitative theory to explain
turnover of interactions in space and time – i.e. a novel approach to
interaction distribution modelling. We apply this novel framework to a large
data set of host–parasite interactions across Europe and find that two aspects
of the environment (temperature and precipitation) leave a strong imprint on
species co-occurrence, but not on the probability of local interactions. Even
where species co-occur, interaction proves a stochastic rather than
deterministic process, adding to variation in realized network structure. We
also find that a large majority of species pairs are never found together,
thus precluding any inferences regarding their probability to interact. Our
framework provides a first conceptual framework to explain the variation of
network structure at large spatial scales and opens new perspectives at the
frontier between species distribution modelling and community ecology.

@@

As a more explicit description of interactions among species, the
network approach offers a convenient representation of communities. Species
are represented as nodes and interactions by links. To date, studies of
network diversity have mostly been concerned with the distribution of
interactions within locations, and less so with the variation among locations
\citep{Dunne2005, Bascompte2007, Ings2007, Kefi2012}. There is however 
ample evidence that interaction networks vary in space and time
\citep{Laliberte2010, Poisot2012, Albouy2014, Poisot2016, Trojelsgaard2015}. Metacommunity
theory provides explanation for the variation in the distribution of the
different nodes \citep{Gravel2011c, Pillai2011}, but there is no explanation
to the joint variation of node and link occurrences. We urgently need a
conceptual framework to formalize these observations.

Given the historically different approaches to modelling the distributions of
species vs. interactions, there is an evident need to bring the two together.
Here, we offer an integrated approach, adopting the view that community
structure is best represented as an ecological network of interactions. Based
on this idea, we propose a new description of the basic concept of the
ecological niche, now integrating the effect of the environment on species
distribution and on the ecological interactions among them. Starting from this
redefined concept, we develop a quantitative theory to explain turnover of
interactions in space and time. We first present the conceptual framework, and
then formalize it mathematically, using a probabilistic model to represent the
sampling of the regional pool of interactions. At the level of species pairs,
the statistical approach could be conceived as an interaction distribution
model. At the community level, the approach provides a likelihood-based method
to compare different hypotheses of network turnover. We apply this novel
framework to a large data set on host–parasite interactions across Europe and
find that variation of the environment causes turnover of both species and
interactions. The network structure changes systematically across the
latitudinal gradient, with a peak of connectance at intermediate latitude.

@@

The problem of community assembly is often formulated as \textit{how are
species sampled from a regional pool to constitute a local community
\citep{Gotzenberger2012}?} This question could be rewritten to address the
problem of network assembly, as \textit{how do samples from a regional pool of
interactions constitute a local interaction network?} An illustration of this
problem for a food web is provided in Figure 1. The metaweb represents
potential interactions among all species that could be found in a given area.
In this particular case, there are 275 nodes, and 1173 links among plants (52
nodes), herbivores (96 nodes), and parasitoids (127 nodes) from Northern
Europe. An instance of a local community is also illustrated, with 45 nodes
and 93 interactions. Only 55.0\% of all potential interactions are realized,
revealing the stochastic nature of ecological interactions. Our objective here
is to provide a conceptual framework to explain the sampling of the regional
pool of interactions, along with a quantitative method to predict it. The
problem could be formalized sequentially by first understanding why only a
fraction of the species are co-occurring locally and second why these species
do or do not interact.

There are multiple causes of spatial turnover in community composition. The
first and most-studied driver is the effect of variation in the abiotic
environment on species performance. Combined with specific responses in
demography, it generates variation among sites by selecting the locally
fittest species \citep{Leibold2004}. Stochasticity plays an additional role,
either colonization and extinction events \citep{Hanski1999} are inherently
unpredictable, or because of strong non-linear feedbacks in community dynamics
generating alternative transients and equilibria \citep{Chase2007,
Vellend2014}. Analyses of community turnover are usually performed with data
represented in a table with rows corresponding to sites (or measurements) and
columns to species. Metrics of beta diversity quantify the variance of this
community data \citep{Legendre2005}. Traditional approaches rely on measures
of dissimilarity among communities, such as the Jaccard or Bray–Curtis
indices. A more recent approach decomposes the total variation of the
community data into species and site contributions to beta diversity
\citep{Legendre2013}, and further partition into dissimilarity due to changes
in species richness and those due to actual species turnover \citep{Baselga2010;
Carvalho et al. 2012}. Even though these methods compare whole lists of
species among sites or measurements, they remain fundamentally ’species-
based’, since they report the variation within columns. None of them
explicitly considers the variation of associations (i.e., of pairs or higher-
order motifs – \citealt{Stouffer2007}).

The "niche" is by far the dominant concept to explain species distributions
and community assembly, from the local to the global scale. Following
\citealt{Hutchinson1957}, the niche is viewed as the set of environmental
conditions allowing a population to establish and persist (see also
\citealt{Holt2009}). Community turnover arises as a result of successive
replacement of species along an environmental gradient, in agreement with the
Gleasonian view of communities \citep{Gleason1926}. The concept is
straightforward to operationalize with species distribution models, as it maps
naturally on available distributional and environmental data. In consequence,
a vast array of statistical tools have been developed to implement it (e.g.
BIOMOD \citealt{Thuiller2003}, MaxEnt \citealt{Phillips2006}). It is however
much harder to account for ecological interactions with this approach
\citep{Peterson2011}. As such, these interactions are often viewed as externalities
constraining or expanding the range of environmental conditions required for a
species to maintain a viable population \citep{Pulliam2000, Soberon2007, Boulangeat2012}.

As mentioned above, in the network approach to community structure, species
and interactions are represented by nodes and links, respectively.
Associations can also be represented by matrices in which entries represent
the occurrence or intensity of interactions among species (rows and columns).
Network complexity is then computed as the number of interactions (in the case
of binary networks) or interaction diversity (in the case of quantitative
networks, \citealt{Bersier2002}). Variability in community structure
consequently arises from the turnover of species composition, along with
turnover of interactions among pairs of species. The occurrence and intensity
of interactions could vary because of the environment, species abundance, and
higher-order interactions \citep{Poisot2015a}. Variation in community
composition was found independent of the variation of ecological interactions,
suggesting that these two components of network variability respond to
different drivers \citep{Poisot2012}.

@@

A more integrative description of the niche is key to understand spatial and
temporal turnover in community structure. Despite several attempts to update
the concept of the ecological niche, ecologists have not moved far past the
``n-dimensional hypervolume'' defined by Hutchinson. Despite its intuitive
interpretation and easy translation into species distribution models
\citep{Boulangeat2012, Blonder2014}, the concept has been frequently
criticized \citep{Hardin1960, Peters1991, Silvertown2004}, and several
attempts have been made to expand and improve it \citep{Pulliam2000,
Chase2003, Soberon2007, Holt2009, McInerny2012b}.

@@

These perspectives are rather orthogonal to each other, which has resulted in
considerable confusion in the literature \citep{McInerny2012a}.
\citealt{Chase2003} attempted to reconcile with the following definition:
``[The niche is] the joint description of the environmental conditions that
allow a species to satisfy its minimum requirements so that the birth rate of
a local population is equal to or greater than its death rate along with the
set of per capita effects of that species on these environmental conditions''.
Their representation merges zero-net growth isoclines delimiting the
Grinnellian niche (when does the population persists) with impact vectors delimiting
the Eltonian niche (what is the per-capita impact). While this representation
has been very influential in local-scale community ecology (the resource-ratio
theory of coexistence, \citealt{Tilman1982}), it remains impractical at larger
spatial scales because of the difficulties in measuring it. The absence of any
mathematical representation of the niche that could easily be fitted to
ecological data may explain why biogeographers are still struggling to develop
species distribution models that also consider ecological interactions.

We propose to integrate the two perspectives of the niche with a visual
representation of both components (Fig. 2). The underlying rationale is that, in
addition to the environmental constraints on demographic performance, any
organism requires resources to sustain its metabolic demands and reproduction.
Abiotic environmental axes are any non-consumable factors affecting the
demographic performance of an organism. Alternatively, the resource axes are
traits of the resources that allow interactions with the consumer. The niche
should therefore be viewed as the set of abiotic environmental factors (the
Grinnellian component) along with the set of traits (the Eltonian component)
that allow a population to establish and to persist at a location. Accordingly,
each species can be characterized by an optimal position in both the
environmental (x-axis) and the trait (y-axis) plane. The integrated niche is
then the hypervolume where interactions can occur and sustain a population.
This approach radically changes the representation of the niche, putting the
distributions and ecological interactions of species in the same formalism.

The limits of the niche axes could be independent of each other (as in the
example in Fig. 2), or  they could interact. For instance, the optimal
prey size for predatory fishes could decline with increasing temperature
\citep{Lelong2015}, which would make diet boundaries functions of the
environment. Alternatively, we could also consider that the growth rate
of the predator changes with the size of its prey items, thereby altering the
environmental boundaries.

@@

We now formalize the integrated niche with a probabilistic approach to
interactions and distributions. We seek to represent the probability that an
interaction between species $i$ and $j$ occurs at location $y$. We define
$L_{ijy}$ as a stochastic variable, and are looking at the probability that
this event occurs, $P(L_{ijy})$. The occurrence of an interaction is dependent
on the co-occurrence of species $i$ and $j$. This argument might seem trivial
at first, but the explicit consideration of this condition in the
probabilistic representation of ecological interactions will prove fundamental
to understand their variation. We define $X_{iy}$ as a stochastic variable
representing the occurrence of species $i$ at location $y$. The quantity we seek to
understand is the probability of a joint event, conditional on the set of
environmental conditions $E_y$:

@@

The first term on the right-hand side of the equation is the probability of
observing the two species co-occurring at location $y$. It corresponds to the
Grinnellian dimension of the niche. The second term represents the probability
that an interaction occurs between species $i$ and $j$, given their set of
traits $T$ and that they are co-occurring. Above, we refer to this entity
as the "metaweb" and it corresponds to the Eltonian dimension of the niche. Below,
we will see how this formalism can be directly fitted to empirical data. But
before turning to an application, we will discuss the interpretation of
different variants of these two terms.

@@

There are several variants to the co-occurrence probability, representing
different hypotheses about temporal and spatial variation in network structure
(see the explicit formulations in Table 1). The simplest model relates the
probability of co-occurrence directly to the environment, $P(X_{iy},X_{jy}|E_y)$.
In this situation there are no underlying assumptions about the ecological
processes responsible for co-occurrence. It could arise because interactions
do constraint distribution \citep{Pollock2014, Cazelles2016} or,
alternatively, because of environmental requirements shared between $i$ and
$j$. In the former case, species are not independent of each other and the
conditional occurrence must be accounted for explicitly, $P(X_{iy},X_{jy}
|E_y)=P(X_{iy}|E_y,X_{jy})P(X_{jy}|E_y)$. In the latter case, species are
independent and only the marginal occurrence must be accounted for, $P(
|E_y)=P(X_{iy} |E_y)P(X_{jy} |E_y)$.

The co-occurrence probability itself could depend on ecological interactions.
This should be viewed as the realized component of the niche (i.e. the
distribution when accounting for species interactions). Direct pairwise
interactions such as competition, facilitation, and predation have long been
studied for their impact on co-distribution (e.g. \citealt{Diamond1976, Connor1980,
Gotelli2000}. Second- and higher-order interactions (e.g. trophic cascades)
could also affect co-occurrence. Co-occurrence of multiple species embedded in
ecological networks is a topic of its own, however, and is influenced by both network
topology and species richness \citep{Cazelles2016}. Not only direct interactions do
influence co-occurrence, but indirect interactions do as well (e.g. plant species
sharing an herbivore could repulse each other in space). The impact of direct
interactions and first-order indirect interactions on co-occurrence tends to
vanish with increasing species richness in the community. Further, co-occurrence
is also influenced by the covariance of interacting species to an
environmental gradient \citep{Cazelles2015}. Because of the complexity of relating
co-occurrence to the structure of interaction networks, we will here focus on
the variation of interactions and not on their distribution, and leave this
specific issue for the Perspectives and future research.

@@

There are also variants of the metaweb. First, most documented metawebs have
thus far considered ecological interactions to be deterministic, not
probabilistic (e.g. \citealt{Havens1992, Wood2015}). Species are assumed to
interact whenever they are found together in a location, independent of their
local abundance and the local environment. In other words,
$P(L_{ijy}|X_{ijy}=1) = 1$ and $P(L_{ijy}|X_{ijy}=0) = 0$. This approach might
be a reasonable approximation if the scale of sampling and inference is so
large that the probability of observing at least one interaction converges to
unity In this cenario, network variation solely arises from species
distributions.

@@

Communities of willows, gallers, and parasitoids are species-rich and
widely distributed, with pronounced variation in community composition across
space. The genus Salix includes over 400 species, most of which are shrubs or
small trees \citep{Argus1997}. The genus is common in most habitats across the
Northern Hemisphere \citep{Skvortsov1999}. Willows support a highly diverse
community of herbivorous insects, with one of the main herbivore groups being
gall-inducing sawflies (Hymenoptera: Tenthredinidae: Nematinae: Euurina
\citep{Kopelke1999}). Gall formation is induced by sawfly females during
oviposition, and includes marked manipulation of host-plant chemistry by the
galler \citep{Nyman2000}. The parasitoid community includes nearly
100 species belonging to 17 insect families of four orders
\citep{Kopelke2000}. These encompass two main types: inquiline larvae
(Coleoptera, Lepidoptera, Diptera, and Hymenoptera) feed primarily on gall
tissue, but typically kill the galler larva in the process, while parasitoid
larvae (representing many families in Hymenoptera) kill the galler larvae by
direct feeding \citep{Kopelke2003}. In terms of associations between the
trophic levels, phylogeny-based comparative studies have demonstrated that
galls represent ``extended phenotypes'' of the gallers, meaning that gall form,
location, and chemistry is determined mainly by the galling insects and not by
their host plants \citep{Nyman2000}. Because galler parasitoids have to
penetrate a protective wall of modified plant tissue in order to gain access
to their victims, gall morphology has been inferred to strongly affect the
associations between parasitoids and hosts \citep{Nyman2007}. Thus, the set of
parasitoids attacking each host is presumably constrained by the form,
size, and thickness of its gall.

Local realizations of the willow–galler–parasitoid network were reconstructed
from community samples collected between 1982 and 2010. During this period,
willow galls were collected at 370 sites across Central and Northern Europe.
Sampling was conducted in the summer months of June and/or July, i.e., during
the later stages of larval development. Galler species were identified on the
basis of willow host species and gall morphology, as these are distinct for
each sawfly species. At each site, galls were randomly collected from numerous
willow individuals in an area of about 0.1–0.3 $km^2$. Some sites were visited
more than once, with a total of 641 site visits across the 370 sites. GPS
coordinates were recorded for each location; for our analyses, the present day
annual mean temperature and precipitation were obtained from WorldClim using
the R package raster \citep{Hijmans2015}. While other covariates could have
also been considered, these variables are likely representative of the most
important axes of the European climate, and more easily interpretable than
reduced variables obtained by e.g. principal component analysis.

@@

We compared three models for the co-occurrence probability. The first one
directly models the co-occurrence probability conditional on the local
environment, $P(X_{iy},X_{jy}|E_y)$ (models are listed at Table 1 and 2).
Hence, this model makes no assumptions about the mechanisms driving co-
occurrence for any given environment, and instead uses the information directly
available in the data. It thereby indirectly accounts for the effect of
interactions on co-occurrence, if there is any. The second model considers
independent occurrence of species. In this case, we independently fit
$P(X_{iy} |E_y)$ and $P(X_{jy} |E_y)$, then we take their product to derive
the probability of co-occurrence. This model should be viewed as a null
hypothesis with respect to the first model, since a comparison between the
respective models will reveal if there is significant spatial association of
the two species beyond a joint response to the shared environment
\citep{Cazelles2015}. Finally, the third model assumes that the probability of
co-occurrence is independent of the environment and thus constant throughout
the landscape. In other words, $P(X_{iy},X_{jy})$ is obtained by simply counting the
number of observed co-occurrences divided by the total number of observations.
Thus, the comparison between the first and third model allows us to test the
hypothesis that co-occurrence is conditional on the environment. Whenever the
environment was included as a covariate in the GLM, we considered a second-
order polynomial response for both temperature and precipitation. There are
consequently 5 parameters for the first model when fitting a given pair of
species, 10 parameters for the second, and only 1 for the third model.

Following the same logic, we compared three models of the interaction
probability. The first model conditions the interaction probability on the
local environmental variables, $P(L_{ijy}|X_{iy},X_{jy},E_y)$. Consequently,
the model was fit to a subset of the data where the two species co-occur.
The second model fits the interaction probability independently of the local
environmental variables, $P(L_{ijy}|X_{iy},X_{jy})$. It corresponds to the
number of times the two species were observed to interact when co-occurring,
divided by the number of times that they co-occurred. The third model is an
extreme case performed only to test the hypothesis that if two species are
found to interact at least once, then they should interact whenever they co-occur,
$P(L_{ijy}|X_{iy},X_{jy})=1$. While not necessarily realistic, this
model tests an assumption commonly invoked in the representation of local
networks from the knowledge of a deterministic metaweb. There are consequently
5 parameters for the first model, a single parameter for the second model and
no parameter to evaluate for the third model (where the interaction
probability is fixed by the hypothesis).

>>>>

10: Centre for Integrative Ecology, School of Biological Sciences, University of Canterbury, Private Bag 4800, Christchurch, New Zealand\\

@@

Biogeography has historically focused on the spatial distribution and
abundance of species, neglecting variation in the way species interact with
each other. Models of species distribution and of interactions have
historically taken different paths. Here, we plea for an integrated approach,
adopting the view that community structure is best represented as an
ecological network of interactions. We outline a conceptual approach,
suggesting that the ecological niche may be redefined to encompass the effect
of the environment on species distribution (the Grinnellian dimension of the
niche) and on the ecological interactions among them (the Eltonian dimension).
Starting from this novel concept, we develop a quantitative theory to explain
turnover of interactions in space and time – i.e. a novel approach to
interaction distribution modelling. We apply this framework to a large
data set of host–parasite interactions across Europe and find that two aspects
of the environment (temperature and precipitation) leave a strong imprint on
species co-occurrence, but not on the probability of local interactions. Even
where species co-occur, interaction proves a stochastic rather than
deterministic process, adding to variation in realized network structure. We
also find that a large majority of species pairs are never found together,
thus precluding any inferences regarding their probability to interact. Our
framework provides a first conceptual framework to explain the variation of
network structure at large spatial scales and opens new perspectives at the
frontier between species distribution modelling and community ecology.

@@

As a more explicit description of interactions among species, the
network approach offers a convenient representation of communities. Species
are represented as nodes and interactions by links. To date, studies of
network diversity have mostly been concerned with the distribution of
interactions within locations, and less so with the variation among locations
\citep{Dunne2005, Bascompte2007, Ings2007, Kefi2012}. There is, however, 
ample evidence that interaction networks vary in space and time
\citep{Laliberte2010, Poisot2012, Albouy2014, Poisot2016, Trojelsgaard2015}. Metacommunity
theory provides an explanation for the variation in the distribution of the
different nodes \citep{Gravel2011c, Pillai2011}, but there is no explanation
to the joint variation of node and link occurrences. Consequently, we urgently need a
conceptual framework to formalize these observations.

Given the historically different approaches to modelling the distributions of
species vs. interactions, there is a clear need to bring the two together.
Here, we offer an integrated approach to do so, adopting the view that community
structure is best represented as an ecological network of interactions. Based
on this idea, we propose a new description of the basic concept of the
ecological niche that integrates the effect of the environment on species
distribution and on the ecological interactions among them. Starting from this
redefined concept, we develop a quantitative theory to explain turnover of
interactions in space and time. We first present the conceptual framework, and
then formalize it mathematically, using a probabilistic model to represent the
sampling of the regional pool of interactions. At the level of species pairs,
the statistical approach could be conceived as an interaction distribution
model. At the community level, the approach provides a likelihood-based method
to compare different hypotheses of network turnover. We apply this novel
framework to a large data set on host–parasite interactions across Europe and
find that variation of the environment causes turnover of both species and
interactions. The network structure changes systematically across the
latitudinal gradient, with a peak of connectance at intermediate latitudes.

@@

The problem of community assembly is often formulated as \textit{how are
species sampled from a regional pool to constitute a local community
\citep{Gotzenberger2012}?} This question could be rewritten to address the
problem of network assembly, as \textit{how do samples from a regional pool of
interactions constitute a local interaction network?} An illustration of this
problem for a food web is provided in Figure 1. The regional pool of interactions, or metaweb, represents
potential interactions among all species that could be found in a given area.
In this particular case, there are 275 nodes, and 1173 links among the plants (52
nodes), herbivores (96 nodes), and parasitoids (127 nodes) from Northern
Europe. An instance of a local community is also illustrated, with 45 nodes
and 93 interactions. Only 55.0\% of all potential interactions are realized locally,
revealing the stochastic nature of ecological interactions. Our objective here
is to provide a conceptual framework to explain the sampling of the regional
pool of interactions, along with a quantitative method to predict it. The
problem could be formalized sequentially by first understanding why only a
fraction of the species are co-occurring locally and second why these species
do or do not interact.

There are multiple causes of spatial turnover in community composition. The
first and most-studied driver is the effect of variation in the abiotic
environment on species performance. Combined with specific responses in
demography, it generates variation among sites by selecting the locally
fittest species \citep{Leibold2004}. Stochasticity plays an additional role,
either because colonization and extinction events \citep{Hanski1999} are inherently
unpredictable or because of strong non-linear feedbacks in community dynamics
generating alternative transients and equilibria \citep{Chase2007,
Vellend2014}. Analyses of community turnover are usually performed with data
represented in a table with rows corresponding to sites (or measurements) and
columns to species. Metrics of beta diversity quantify the variance of this
community data \citep{Legendre2005}. Traditional approaches rely on measures
of dissimilarity among communities, such as the Jaccard or Bray–Curtis
indices. A more recent approach decomposes the total variation of the
community data into species and site contributions to beta diversity
\citep{Legendre2013}, and further partitions into dissimilarity due to changes
in species richness and those due to actual species turnover \citep{Baselga2010;
Carvalho et al. 2012}. Even though these methods compare whole lists of
species among sites or measurements, they remain fundamentally `species-based',
since they report the variation within columns. None of them
explicitly considers the variation of associations (i.e., of pairs or higher-order
motifs – \citealt{Stouffer2007}).

The ``niche'' is by far the dominant concept to explain species distributions
and community assembly, from the local to the global scale. Following
\citealt{Hutchinson1957}, the niche is viewed as the set of environmental
conditions allowing a population to establish and persist (see also
\citealt{Holt2009}). Community turnover arises as a result of successive
replacement of species along an environmental gradient, in agreement with the
Gleasonian view of communities \citep{Gleason1926}. The concept is
straightforward to operationalize with species distribution models, as it maps
naturally on available distributional and environmental data. In consequence,
a vast array of statistical tools have been developed to implement it (e.g.
BIOMOD \citealt{Thuiller2003}, MaxEnt \citealt{Phillips2006}). It is however
much harder to account for ecological interactions within this approach
\citep{Peterson2011}. As such, these interactions are often viewed as externalities
constraining or expanding the range of environmental conditions required for a
species to maintain a viable population \citep{Pulliam2000, Soberon2007, Boulangeat2012}.

As mentioned above, in the network approach to community structure, species
and interactions are represented by nodes and links, respectively.
Associations can also be represented by matrices in which entries represent
the occurrence or intensity of interactions among species (rows and columns).
Network complexity is then computed as the number of interactions (in the case
of binary networks) or interaction diversity (in the case of quantitative
networks, \citealt{Bersier2002}). Variability in community structure
consequently arises from the turnover of species composition, along with
turnover of interactions among pairs of species. The occurrence and intensity
of interactions could vary because of the environment, species abundance, and
higher-order interactions \citep{Poisot2015a}. Variation in community
composition was found to be independent of the variation of ecological interactions,
suggesting that these two components of network variability may well respond to
different drivers \citep{Poisot2012}.

@@

A more integrative description of the niche is key to understanding spatial and
temporal turnover in community structure. Despite several attempts to update
the concept of the ecological niche, ecologists have not moved far beyond the
``n-dimensional hypervolume'' defined by Hutchinson. Despite its intuitive
interpretation and easy translation into species distribution models
\citep{Boulangeat2012, Blonder2014}, the concept has been frequently
criticized \citep{Hardin1960, Peters1991, Silvertown2004}, and several
attempts have been made to expand and improve it \citep{Pulliam2000,
Chase2003, Soberon2007, Holt2009, McInerny2012b}.

@@

These perspectives are rather orthogonal to each other, and this has resulted in
considerable confusion in the literature \citep{McInerny2012a}.
\citealt{Chase2003} attempted to reconcile with the following definition:
``[The niche is] the joint description of the environmental conditions that
allow a species to satisfy its minimum requirements so that the birth rate of
a local population is equal to or greater than its death rate along with the
set of per capita effects of that species on these environmental conditions''.
Their representation merges zero-net growth isoclines delimiting the
Grinnellian niche (when does the population persists) with impact vectors delimiting
the Eltonian niche (what is the per-capita impact). While this representation
has been very influential in local-scale community ecology (the resource-ratio
theory of coexistence, \citealt{Tilman1982}), it remains impractical at larger
spatial scales because of the difficulties in measuring it. The absence of any
mathematical representation of the niche that could be easily fit to
ecological data may explain why biogeographers are still struggling to develop
species distribution models that also consider ecological interactions.

We propose to integrate the two perspectives of the niche with a visual
representation of both components (Fig. 2). The underlying rationale is that, in
addition to the environmental constraints on demographic performance, any
organism requires resources to sustain its metabolic demands and reproduction.
Abiotic environmental axes are any non-consumable factors affecting the
demographic performance of an organism. Alternatively, the resource axes are
traits of the resources that allow interactions with the consumer. The niche
should therefore be viewed as the set of abiotic environmental factors (the
Grinnellian component) along with the set of traits (the Eltonian component)
that allow a population to establish and to persist at a location. Accordingly,
each species can be characterized by an optimal position in both the
environmental (x-axis) and the trait (y-axis) plane. The integrated niche is
then the hypervolume where interactions can occur and sustain a population.

This approach radically changes the representation of the niche, putting species
distributions and ecological interactions into the same formalism.
Moreover, it allows the limits of the niche axes to be independent of each other (as in the
example in Fig. 2), or to interact. For instance, the optimal
prey size for predatory fishes could decline with increasing temperature
\citep{Lelong2015}, which would make diet boundaries functions of the
environment. Alternatively, we could also consider that the growth rate
of the predator changes with the size of its prey items, thereby altering the
environmental boundaries.

@@

We now formalize the integrated niche with a probabilistic approach to
interactions and distributions. In particular, we seek to represent the probability that an
interaction between species $i$ and $j$ occurs at location $y$. We define
$L_{ijy}$ as a stochastic variable, and focus on the probability that
this event occurs, $P(L_{ijy})$. We note that the occurrence of an interaction is dependent
on the co-occurrence of species $i$ and $j$. This argument might seem trivial
at first, but the explicit consideration of this condition in the
probabilistic representation of ecological interactions will prove instrumental
to understand their variation. We define $X_{iy}$ as a stochastic variable
representing the occurrence of species $i$ at location $y$. The quantity we seek to
understand is the probability of a joint event, conditional on the set of
environmental conditions $E_y$:

@@

The first term on the right-hand side of the equation is the probability of
observing the two species co-occurring at location $y$. It corresponds to the
Grinnellian dimension of the niche. The second term represents the probability
that an interaction occurs between species $i$ and $j$, given their set of
traits $T$ and that they are co-occurring. Above, we referred to this entity
as the ``metaweb'' and it corresponds to the Eltonian dimension of the niche. Below,
we will see how this formalism can be directly fit to empirical data. But
before turning to an application, we will discuss the interpretation of
different variants of these two terms.

@@

There are several variants to the co-occurrence probability, representing
different hypotheses about temporal and spatial variation in network structure
(see the explicit formulations in Table 1). The simplest model relates the
probability of co-occurrence directly to the environment, $P(X_{iy},X_{jy}|E_y)$.
In this situation, there are no underlying assumptions about the ecological
processes responsible for co-occurrence. It could arise because interactions
do in fact constrain distributions \citep{Pollock2014, Cazelles2016} or,
alternatively, because of environmental requirements shared between $i$ and
$j$. In the former case, species are not independent of each other and the
conditional occurrence must be accounted for explicitly, $P(X_{iy},X_{jy}
|E_y)=P(X_{iy}|E_y,X_{jy})P(X_{jy}|E_y)$. In the latter case, species are
independent, and only the marginal occurrence must be accounted for, $P(
|E_y)=P(X_{iy} |E_y)P(X_{jy} |E_y)$.

The co-occurrence probability itself could depend on ecological interactions.
This should be viewed as the realized component of the niche (i.e. the
distribution when accounting for species interactions). Direct pairwise
interactions such as competition, facilitation, and predation have long been
studied for their impact on co-distribution (e.g. \citealt{Diamond1976, Connor1980,
Gotelli2000}). Second- and higher-order interactions (e.g. trophic cascades)
could also affect co-occurrence. Co-occurrence of multiple species embedded in
ecological networks is a topic of its own, however, and is influenced by both network
topology and species richness \citep{Cazelles2016}. Not only do direct interactions
influence co-occurrence, but indirect interactions do as well (e.g. plant species
sharing an herbivore could repel each other in space). The impact of direct
interactions and first-order indirect interactions on co-occurrence tends to
vanish with increasing species richness in the community. Further, co-occurrence
is also influenced by the covariance of interacting species to an
environmental gradient \citep{Cazelles2015}. Because of the complexity of relating
co-occurrence to the structure of interaction networks, we will focus here on
the variation of interactions and not on their distribution, and leave this
specific issue for the Perspectives and future research.

@@

There are also variants of the metaweb. First, most documented metawebs have
thus far considered ecological interactions to be deterministic, not
probabilistic (e.g. \citealt{Havens1992, Wood2015}). Species are assumed to
interact whenever they are found together in a location, independent of their
local abundance and the local environment. In other words,
$P(L_{ijy}|X_{ijy}=1) = 1$ and $P(L_{ijy}|X_{ijy}=0) = 0$. This approach might
be a reasonable approximation if the scale of sampling and inference is so
large that the probability of observing at least one interaction converges to
unity. In this scenario, network variation solely arises from species
distributions.

@@

Communities of willows, gallers, and parasitoids are species-rich and
widely distributed, with pronounced variation in community composition across
space. The genus Salix includes over 400 species, most of which are shrubs or
small trees \citep{Argus1997}, and is common in most habitats across the
Northern Hemisphere \citep{Skvortsov1999}. Willows support a highly diverse
community of herbivorous insects, with one of the main herbivore groups being
gall-inducing sawflies (Hymenoptera: Tenthredinidae: Nematinae: Euurina
\citep{Kopelke1999}). Gall formation is induced by sawfly females during
oviposition, and includes marked manipulation of host-plant chemistry by the
galler \citep{Nyman2000}. The parasitoid community includes nearly
100 species belonging to 17 insect families of four orders
\citep{Kopelke2000}. These encompass two main types: inquiline larvae
(Coleoptera, Lepidoptera, Diptera, and Hymenoptera) feed primarily on gall
tissue, but typically kill the galler larva in the process, while parasitoid
larvae (representing many families in Hymenoptera) kill the galler larvae by
direct feeding \citep{Kopelke2003}. In terms of associations between the
trophic levels, phylogeny-based comparative studies have demonstrated that
galls represent ``extended phenotypes'' of the gallers, meaning that gall form,
location, and chemistry is determined mainly by the galling insects and not by
their host plants \citep{Nyman2000}. Because galler parasitoids have to
penetrate a protective wall of modified plant tissue in order to gain access
to their victims, gall morphology has been inferred to strongly affect the
associations between parasitoids and hosts \citep{Nyman2007}. Thus, the set of
parasitoids attacking each host is presumably constrained by the form,
size, and thickness of its gall.

Local realizations of the willow–galler–parasitoid network were reconstructed
from community samples collected between 1982 and 2010. During this period,
willow galls were collected at 370 sites across Central and Northern Europe.
Sampling was conducted in the summer months of June and/or July, i.e., during
the later stages of larval development. Galler species were identified on the
basis of willow host species and gall morphology, as these are distinct for
each sawfly species. At each site, galls were randomly collected from numerous
willow individuals in an area of about 0.1–0.3 $km^2$. Some sites were visited
more than once, with a total of 641 site visits across the 370 sites. GPS
coordinates were recorded for each location; for our analyses, the present day
annual mean temperature and precipitation were obtained from WorldClim using
the R package raster \citep{Hijmans2015}. While other covariates could have
also been considered, these variables are likely representative of the most
important axes of the European climate, and more easily interpretable than
reduced variables obtained, for example, by principal component analysis.

@@

We compared three models for the co-occurrence probability. The first one
directly models the co-occurrence probability conditional on the local
environment, $P(X_{iy},X_{jy}|E_y)$ (models are listed at Table 1 and 2).
Hence, this model makes no assumptions about the mechanisms driving co-occurrence
for any given environment, and instead uses the information directly
available in the data. It thereby indirectly accounts for the effect of
interactions on co-occurrence, if there are any. The second model considers
independent occurrence of species. In this case, we independently fit
$P(X_{iy} |E_y)$ and $P(X_{jy} |E_y)$, and we then take their product to derive
the probability of co-occurrence. This model should be viewed as a null
hypothesis with respect to the first model, since a comparison between the
respective models will reveal if there is significant spatial association of
the two species beyond a joint response to the shared environment
\citep{Cazelles2015}. Finally, the third model assumes that the probability of
co-occurrence is independent of the environment and thus constant throughout
the landscape. In other words, $P(X_{iy},X_{jy})$ is obtained by simply counting the
number of observed co-occurrences divided by the total number of observations.
Thus, the comparison between the first and third model allows us to test the
hypothesis that co-occurrence is conditional on the environment. Whenever the
environment was included as a covariate in the GLM, we considered a second-order
polynomial response for both temperature and precipitation. There are
consequently 5 parameters for the first model when fitting a given pair of
species, 10 parameters for the second, and only 1 for the third model.

Following the same logic, we compared three models of the interaction
probability. The first model conditions the interaction probability on the
local environmental variables, $P(L_{ijy}|X_{iy},X_{jy},E_y)$. Consequently,
the model was fit to the subset of the data where the two species co-occur.
The second model fits the interaction probability independently of the local
environmental variables, $P(L_{ijy}|X_{iy},X_{jy})$. It corresponds to the
number of times the two species were observed to interact when co-occurring,
divided by the number of times that they co-occurred. The third model is an
extreme case performed only to test the hypothesis that if two species are
found to interact at least once, then they should interact whenever they co-occur,
$P(L_{ijy}|X_{iy},X_{jy})=1$. While not necessarily realistic, this
model tests an assumption commonly invoked in the representation of local
networks from the knowledge of a deterministic metaweb. There are consequently
5 parameters for the first model, a single parameter for the second model and
no parameter to evaluate for the third model (where the interaction
probability is fixed by the hypothesis).
